Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame

### 2022-09-19 manuscript/Chapter3-manuscript.Rmd
added paragraph describind study system. 1st draft

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging. Importantly, we have little information on the ecological contexts - except for competition - that favors one strategy over the other. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specializing [@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. While quantifying the fitness consequences of specialization is a daunting task, in predator-prey systems, some clues indicate that fluctuations in the predictability of prey (e.g. abundance, availability, behavior) encounters throughout a predator's lifetime may be a key factor [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability hypothesis advances that when resources are predictable, individual specialists should have higher delivery rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should be advantaged when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019].

In this study, we analyze individual behavioral data from players in *Dead by Daylight* to test how hunting expertise and prey behavior shape predator foraging specialization. *Dead by Daylight* is an online videogame simulating a predator-prey interaction, where one predator player hunts four prey players in different virtual environments. The data grants a high degree of precision on the behavioral interaction as both the predator and the four prey were measured simultaneously in each trial. First, we investigate how predators developed their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise because they encounter varying levels of difficulty with the prey that they hunt. Second, we test the hypothesis that individual variation in foraging behavior will change with expertise. If expertise reduces the costs of switching between hunting tactics, we predict that individual predators should become more flexible with time. Alternatively, if expertise enables the refinement of the hunting tactics, then individuals may instead specialize. Third, we evaluate how predator foraging specialization interacts with variation in prey behavior. Whether predators specialize or not with expertise may depend on the behavior of their prey, for instance, as it may be harder to specialize when encounters are less predictable. Lastly, if individuals differ in their degree of foraging specialization, then we expect that the success of specialists and generalists will be equal. Specialist hunters should fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.

@@

>>>>

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging. Importantly, we have little information on the ecological contexts - except for competition - that favor one strategy over the other. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specializing [@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. While quantifying the fitness consequences of specialization is a daunting task, in predator-prey systems, some clues indicate that fluctuations in the predictability of prey encounters throughout a predator's lifetime may be a key factor [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability hypothesis advances that when resources are predictable, individual specialists should have higher delivery rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should be advantaged when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019]. This hypothesis has however not been tested in systems where behavioral interactions are simultaneously monitored between individuals.

In this study, we analyze individual behavioral data from players in *Dead by Daylight* to test how hunting expertise and prey behavior shape predator foraging specialization. *Dead by Daylight* is an online videogame simulating a predator-prey interaction, where one predator player hunts four prey players in different virtual environments. The data grants a high degree of precision on the behavioral interaction as both the predator and the four prey were measured simultaneously in each trial. First, we investigate how predators developed their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise because they encounter varying levels of difficulty with the prey that they hunt. Second, we test the hypothesis that individual variation in foraging behavior will change with expertise. If expertise reduces the costs of switching between hunting tactics, we predict that individual predators should become more flexible with time. Alternatively, if expertise enables the refinement of the hunting tactics, then individuals may instead specialize. Third, we evaluate how predator foraging specialization interacts with variation in prey behavior. Whether predators specialize or not with expertise may depend on the behavior of their prey, for instance, as it may be harder to specialize when encounters are less predictable. Lastly, if individuals differ in their degree of foraging specialization, then we expect specialist hunters to fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.

@@

*Dead by Daylight* is a survival horror asymmetric multiplayer online game developed by the videogame company Behavior Interactive Inc. In this game, players get to chose if they wish to play has a predator or a prey. The objective of the predator player is to search, stalk, and hunt the four prey players across a virtual environment. The objective of the four prey players is to forage for resources in the form of generators that once activated, will enable them to escape through an exit door and win the match. Each player, predator or prey, can chose an avatar with abilities that encourage specific play styles (e.g. bolder vs cautious prey, or ambush vs roaming predator). During our study period, the game offered X playable predator avatars, and X prey avatars. The virtual environment where matches take place is composed of fixed and procedurally generated items (e.g. vegetation, mazes, buildings) that constitute its structure. Some of these environments are larger than others, with varying levels or buildings. However, we have previously shown that they generally do not influence the behavior nor hunting success of the predator, which is probably due to a feature where predators have visual cues on the generators [@FraserFranco.etal2022].

Game objectives and what prey and predators can do...
